
1905 project code, Robert Orpet, 22 June, 2022
For: Reflective groundcovers implementation manuscript

Design: complete block with two treatments (control, reflective fabric (= Extenday). Blocks are adjacent plots.

Sampling: there were pre-treatment and post-treatment samples. Each sample included multiple measurements (beat tray taps) on the same plot. There were multiple post-treatment samples through spring (beat tray taps, spur counts, and leaf counts).

Analysis approach: summarize multiple measurements to one mean per plot per date

```{r SETUP}
#LOAD PACKAGES
library(readxl)
library(ggplot2)
library(ggpubr)
library(emmeans)


#LOAD DATA
Data_taps <- read_excel("1905 - Data - Adults Taps.xlsx", skip = 9)
Data_spurs <- read_excel("1905 - Data - Spurs.xlsx", skip = 9)
Data_leaves <- read_excel("1905 - Data - Leaves.xlsx", skip = 9)

#RELABEL TREATMENT NAMES 
Data_taps$trt <- factor(Data_taps$trt,
                    levels=c("A","B"),
                    labels=c("Control", "Extenday"))
Data_spurs$trt <- factor(Data_spurs$trt,
                    levels=c("A","B"),
                    labels=c("Control", "Extenday"))
Data_leaves$trt <- factor(Data_leaves$trt,
                    levels=c("A","B"),
                    labels=c("Control", "Extenday"))

#CREATE COLUMN FOR DATES AS FACTORS
Data_taps$date_factor <- factor(Data_taps$date)
Data_spurs$date_factor <- factor(Data_spurs$date)
Data_leaves$date_factor <- factor(Data_leaves$date)

#SUMMARIZE MULTIPLE MEASUREMENTS TO ONE MEAN PER PLOT PER DATE
Data_taps <- do.call(data.frame, 
             aggregate(adults~rep+trt+date+date_factor, 
                       data = Data_taps, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
Data_spurs <- do.call(data.frame, 
             aggregate(cbind(eggs,nymphs)~rep+trt+date, 
                       data = Data_spurs, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
Data_leaves <- do.call(data.frame, 
             aggregate(cbind(eggs,yng,old)~rep+trt+date, 
                       data = Data_leaves, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))

#GRAPH APPEARANCE
theme_new <- theme_set(theme_bw())
theme_new <- theme_update(
  plot.background = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.title.x = element_blank(),
  axis.text.x = element_text(color = 'black'),
  axis.text.y = element_text(color = 'black'),
  legend.position=c(1,1),
  legend.justification=c(1,0),
  legend.box="vertical",
  legend.direction='vertical',
  legend.box.margin=margin(0,0,0,0),
  legend.title = element_blank(),
  legend.key.size = unit(0.75, 'lines')
)

```

```{r VISUALIZATION}
#extenday was installed april 3, removal may 13

adults <- ggplot(Data_taps, aes(x = date, y = adults.mean, fill = trt, shape = trt, linetype = trt, color = trt)) +
  scale_fill_manual(values = c('black', 'white')) +
  scale_color_manual(values = c('black', 'grey')) +
  scale_shape_manual(values = c(21,23)) +
  geom_vline(xintercept = as.POSIXct(as.Date("2019-04-03")), linetype=2) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0, linetype = 'solid') +
  stat_summary(fun = "mean", geom = "point") +
  #scale_x_continuous("", limits = c(59,135), breaks = c(60,91,121), labels = c("1-Mar","1-Apr","1-May")) +
  scale_y_continuous(limits = c(0,NA), "No. adults/tray")

eggs <- ggplot(Data_spurs, aes(x = date, y = eggs.mean, fill = trt, shape = trt, linetype = trt, color = trt)) +
  scale_fill_manual(values = c('black', 'white')) +
  scale_color_manual(values = c('black', 'grey')) +
  scale_shape_manual(values = c(21,23,2)) +
  geom_vline(xintercept = as.POSIXct(as.Date("2019-04-03")), linetype=2) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0, linetype = 'solid') +
  stat_summary(fun = "mean", geom = "point") +
  #scale_x_continuous("", limits = c(59,135), breaks = c(60,91,121), labels = c("1-Mar","1-Apr","1-May")) +
  scale_y_continuous(limits = c(0,NA), "Average eggs (buds)")

nymphs <- ggplot(Data_spurs, aes(x = date, y = nymphs.mean, fill = trt, shape = trt, linetype = trt, color = trt)) +
  scale_fill_manual(values = c('black', 'white')) +
  scale_color_manual(values = c('black', 'grey')) +
  scale_shape_manual(values = c(21,23)) +
  geom_vline(xintercept = as.POSIXct(as.Date("2019-04-03")), linetype=2) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0, linetype = 'solid') +
  stat_summary(fun = "mean", geom = "point") +
  #scale_x_continuous("", limits = c(59,135), breaks = c(60,91,121), labels = c("1-Mar","1-Apr","1-May")) +
  scale_y_continuous(limits = c(0,NA),"Average nymphs (buds)")

eggs_leaves <- ggplot(Data_leaves, aes(x = date, y = eggs.mean, fill = trt, shape = trt, linetype = trt, color = trt)) +
  scale_fill_manual(values = c('black', 'white')) +
  scale_color_manual(values = c('black', 'grey')) +
  scale_shape_manual(values = c(21,23,2)) +
  geom_vline(xintercept = as.POSIXct(as.Date("2019-04-03")), linetype=2) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0, linetype = 'solid') +
  stat_summary(fun = "mean", geom = "point") +
  #scale_x_continuous("", limits = c(59,135), breaks = c(60,91,121), labels = c("1-Mar","1-Apr","1-May")) +
  scale_y_continuous(limits = c(0,NA),"Average eggs (leaves)")

nymphs_leaves <- ggplot(Data_leaves, aes(x = date, y = yng.mean+old.mean, fill = trt, shape = trt, linetype = trt, color = trt)) +
  scale_fill_manual(values = c('black', 'white')) +
  scale_color_manual(values = c('black', 'grey')) +
  scale_shape_manual(values = c(21,23,2)) +
  geom_vline(xintercept = as.POSIXct(as.Date("2019-04-03")), linetype=2) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0, linetype = 'solid') +
  stat_summary(fun = "mean", geom = "point") +
  #scale_x_continuous("", limits = c(59,135), breaks = c(60,91,121), labels = c("1-Mar","1-Apr","1-May")) +
  scale_y_continuous(limits = c(0,NA),"Average nymphs (leaves)")

```

```{r}
#STATISTICAL ANALYSIS

#adults taps - pre-installation: P = 0.51
data_test <- subset(Data_taps, date_factor == "2019-04-01")
model <- aov(adults.mean ~ trt + rep, data = data_test)
summary(model)
table_summary <- do.call(data.frame, 
             aggregate(adults.mean~trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                         SEM = sd(x)/(sqrt(length(x))), 
                       count = length(x))))
  
#adults taps - post-installation: P = 0.085
data_test <- subset(Data_taps, date_factor != "2019-04-01")
data_test <- do.call(data.frame, 
             aggregate(adults.mean~rep+trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                       count = length(x))))
model <- aov(adults.mean.mean ~ trt + rep, data = data_test)
summary(model)
table_summary <- do.call(data.frame, 
             aggregate(adults.mean.mean~trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                         SEM = sd(x)/(sqrt(length(x))), 
                       count = length(x))))

#eggs (spurs): P = 0.040
data_test <- do.call(data.frame, 
             aggregate(eggs.mean~rep+trt, 
                       data = Data_spurs, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
model <- aov(eggs.mean.mean ~ trt + rep, data = data_test)
summary(model)
table_summary <- do.call(data.frame, 
             aggregate(eggs.mean.mean~trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                         SEM = sd(x)/(sqrt(length(x))), 
                       count = length(x))))

#nymphs (spurs): P = 0.45
data_test <- do.call(data.frame, 
             aggregate(nymphs.mean~rep+trt, 
                       data = Data_spurs, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
model <- aov(nymphs.mean.mean ~ trt + rep, data = data_test)
summary(model)
table_summary <- do.call(data.frame, 
             aggregate(nymphs.mean.mean~trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                         SEM = sd(x)/(sqrt(length(x))), 
                       count = length(x))))

#nymphs (leaves): P = 0.16
Data_leaves$nymphs <- Data_leaves$yng.mean + Data_leaves$old.mean
data_test <- do.call(data.frame, 
             aggregate(nymphs~rep+trt, 
                       data = Data_leaves, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
model <- aov(nymphs.mean ~ trt + rep, data = data_test)
summary(model)
table_summary <- do.call(data.frame, 
             aggregate(nymphs.mean~trt, 
                       data = data_test, 
                       function(x) 
                       c(mean = mean(x),
                         SEM = sd(x)/(sqrt(length(x))), 
                       count = length(x))))

```
