
2003 project code, Robert Orpet, 8 April, 2022
For: Reflective groundcovers implementation manuscript

Design: randomized complete block with three treatments (control, reflective fabric (= Extenday), metallized film (= mylar)). Blocks are pear orchards (referred to as block or site).

Sampling: there were pre-treatment and post-treatment samples. Each sample included multiple measurements on the same plot. There were multiple post-treatment samples through spring. In summer, cooperators removed treatments at variable timings, so analysis focuses on spring.

Analysis approach: analyses were repeated with multiple approaches for comparison because there are multiple philosophies on use of random effects and mixed models.
Method 1. summarize repeated measures within plots with average, summarize repeated measures across time post-treatment with cumulative insect-days per day, use random effect for blocks. Uses glmmTMB()
Method 2. as method 1, but uses lme()
Method 3. as method 1, but fixed effect for block and uses aov()
Method 4. no summarizing, use random effects for plots nested within block, and use negative binomial distribution. Uses glmmTMB()


```{r Chunk 1 Setup}

#Load packages
library(readxl)
library(ggplot2)
library(ggpubr)
library(glmmTMB)
library(DHARMa)
library(car)
library(emmeans)
library(nlme)

#Load data
Data <- read_excel("2003 - data - beats and buds.xlsx",
                   skip = 9,
                   guess_max = 10000)

#Data cleanup and labeling

#there were two pre-treatment samples at site E instead of one. Exclude the first one so we only visualize and analyze the one sample closer to groundcover installation
Data <- subset(Data, !(Site == "E" & day == 77))
#Eggs were counted on wood and green parts of buds. Calculate sum of eggs for the two locations on buds for analysis.
Data$budcount_eggs <- Data$Eggs.bud + Data$Eggs.wood
#Subset by season for future analyses
Data_spring_all <- subset(Data, Phase != 'Leaves') #data up to summer when switch from bud to leaf monitoring occured
#Rename site labels for visualizations
Data_spring_all$Site <- as.factor(Data_spring_all$Site)
levels(Data_spring_all$Site)[levels(Data_spring_all$Site)=="A"] <- "Site A"
levels(Data_spring_all$Site)[levels(Data_spring_all$Site)=="B"] <- "Site B"
levels(Data_spring_all$Site)[levels(Data_spring_all$Site)=="C"] <- "Site C"
levels(Data_spring_all$Site)[levels(Data_spring_all$Site)=="D"] <- "Site D"
levels(Data_spring_all$Site)[levels(Data_spring_all$Site)=="E"] <- "Site E"
#Rename treatment labels for visualizations
Data_spring_all$Trt <- as.factor(Data_spring_all$Trt)
levels(Data_spring_all$Trt)[levels(Data_spring_all$Trt)=="Control"] <- "Control"
levels(Data_spring_all$Trt)[levels(Data_spring_all$Trt)=="Mylar"] <- "Metallized film"
levels(Data_spring_all$Trt)[levels(Data_spring_all$Trt)=="Extenday"] <- "Reflective fabric"

```

```{r Chunk 2 Visualization: Fig 2. Pear psylla adults and eggs pre-installation through spring}

#Theme for graphs
theme_set(theme_classic())
  theme_update(
        strip.text = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        axis.line = element_line(colour = 'black', size = 0),
        axis.ticks = element_line(colour = "black"),
        axis.text.y = element_text(color = 'black'),
        axis.text.x = element_text(color = "black"),
        panel.spacing = unit(0.5, "lines"),
  legend.position=c(1,1),
  legend.justification=c(1,0),
  legend.box="vertical",
  legend.direction='vertical',
  legend.box.margin=margin(-10,0,5,0), #top,right,bottom,left
  legend.title = element_blank(),
  legend.key.size = unit(0.75, 'lines')
        )

#Springtime graphs of adults [Figure 2A]
adults <- ggplot(data = Data_spring_all, aes(x = day, y = Adults, shape = Trt, fill = Trt, color = Trt)) +
  geom_hline(data = subset(Data_spring_all, Site == "Site E"), aes(yintercept = 8), color = 'white') + #this expands the y axis a little to prevent decimal digits in that panel
  scale_fill_manual(values = c('black', 'white', 'grey')) +
  scale_color_manual(values = c('black', 'grey', 'black')) +
  scale_shape_manual(values = c(21,23,23)) +
  scale_y_continuous(limits = c(0, NA)) + 
  stat_summary(fun.y = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0) +
  stat_summary(fun.y = mean, geom = "point") +
  facet_wrap(~Site, scales = "free_y", nrow = 1) +
  #Vertical lines for DEPLOYMENT
  geom_vline(data = subset(Data_spring_all, Site == "Site A"), aes(xintercept = 78)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site B"), aes(xintercept = 82)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site C"), aes(xintercept = 81)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site D"), aes(xintercept = 82)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site E"), aes(xintercept = 92)) +
  scale_x_continuous(breaks=c(80, 92, 106),labels=c("20\nMar", "1\nApr", "15\nApr")) +
  ylab('No. adults') +
  xlab('') +
  guides(shape = guide_legend(nrow = 3))  +
  theme(
  strip.text = element_text(size = 10, hjust = 0),
  plot.margin=unit(c(-0.2,0.1,-1,0.1),"cm")
  )

#Springtime graphs of adults [Figure 2B]
eggs <- ggplot(data = Data_spring_all, aes(x = day, y = Eggs.bud+Eggs.wood, shape = Trt, fill = Trt, color = Trt)) +
  geom_hline(data = subset(Data_spring_all, Site == "Site D"), aes(yintercept = 13), color = 'white') + #this expands the y axis a little to prevent decimal digits in that panel
  scale_fill_manual(values = c('black', 'white', 'grey')) +
  scale_color_manual(values = c('black', 'grey', 'black')) +
  scale_shape_manual(values = c(21,23,23)) +
  scale_y_continuous(limits = c(0, NA)) + 
  stat_summary(fun.y = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0) +
  stat_summary(fun.y = mean, geom = "point") +
  facet_wrap(~Site, scales = "free_y", nrow = 1) +
  #Vertical lines for DEPLOYMENT
  geom_vline(data = subset(Data_spring_all, Site == "Site A"), aes(xintercept = 78)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site B"), aes(xintercept = 82)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site C"), aes(xintercept = 81)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site D"), aes(xintercept = 82)) +
  geom_vline(data = subset(Data_spring_all, Site == "Site E"), aes(xintercept = 92)) +
  scale_x_continuous(breaks=c(80, 92, 106),labels=c("20\nMar", "1\nApr", "15\nApr")) +
  ylab('No. eggs') +
  xlab('') +
  guides(shape = guide_legend(nrow = 3)) +
  theme(plot.margin=unit(c(-1,0.1,-0.2,0.1),"cm"))

#output
Fig2 <- (ggarrange(
    adults,
    eggs,
    labels = c("A","B"), vjust = c(-0.5,-0.5),
    ncol = 1,
    align="hv",common.legend = TRUE
    ))
ggsave(plot = Fig2, file = "Fig2.png", width = 17, height = 8, unit = "cm")

```

```{r Chunk 3 Summarization: Calculate pear psylla adult insect-days}

#approach: treat each of the five sites as a block, so N = 5
#calculate cumulative insect days for each life stage by site

#First, summarize the multiple counts per plot into one number per plot per site per visit
Data_adults <- do.call(data.frame, 
             aggregate(Adults~Phase+Date+day+Site+Trt, 
                       data = Data, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))

#Calculate cumulative insect days - adult taps
#Build blank data frame
CID_output <- data.frame(Season=character(), 
                         Block=character(), 
                         Treatment=character(), 
                         CID=numeric(), 
                         Days_sum=numeric())
Data_adults$ID <- paste(Data_adults$Phase,Data_adults$Site,Data_adults$Trt,sep = "_")  #Make a unique identifer for each experimental unit by Period of study
IDs <- unique(Data_adults$ID)                                   #Get a list of the IDs
for (value in IDs) {                                            #loop for each ID
  Data_subset <- subset(Data_adults, ID == value)
  row.names(Data_subset) <- NULL
  for (row in 1:nrow(Data_subset)) {                          #loop to calculate within each ID
    if(row == 1){
      Insect_avg <- NA
      Days_difference <- NA
      Insectdays <- 0
      Days <- 0
      output <- data.frame(Insectdays,Days)
    }
    #pre-sample was only one date, so use the point data instead of trying to calculate CIDs
    if(unique(Data_subset$Phase == "Pretreatment")){
      Days <- 1
      Insectdays <- Data_subset[row,"Adults.mean"]
      output <- data.frame(Insectdays,Days)
    }
    if(row > 1){
    Insect_current <- Data_subset[row,"Adults.mean"]
    Insect_previous <- Data_subset[(row-1),"Adults.mean"]
    Insect_avg <- (Insect_current + Insect_previous) / 2
    Day_current <- as.numeric(Data_subset[row,"day"])
    Day_previous <- as.numeric(Data_subset[row-1,"day"])
    if(Day_current < Day_previous){                             #check for out-of-order dates
      print("Error: Days not sorted chronologically")
      print(value)
    }
    Days <- Day_current - Day_previous
    Insectdays <- Insect_avg*Days
    newrow <- data.frame(Insectdays,Days)
    output <- rbind(output, newrow)
    }
  }
  Season <- unique(Data_subset$Phase)
  Block <- unique(Data_subset$Site)
  Treatment <- unique(Data_subset$Trt)
  CID <- sum(output$Insectdays)
  Days_sum <- sum(output$Days)
  CID_newrow <- data.frame(Season,Block,Treatment, CID, Days_sum)
  CID_output <- rbind(CID_output, CID_newrow)
}

#Don't want want to analyze summer data due to reflective materials being removed at different times between sites
CID_output <- subset(CID_output, Season != 'Leaves')
CID_output$Season <- factor(CID_output$Season,
                    levels=c("Pretreatment","Buds"),
                    labels=c("Pre-installation", "Spring"))
CID_output$Treatment <- factor(CID_output$Treatment,
                    levels=c("Control","Extenday","Mylar"),
                    labels=c("Ctrl", "R. fab.","M. film"))

#Adults seasonal mean calculation
CID_output$seasonalmean <- (CID_output$CID / CID_output$Days_sum)
CID_output_adults <- CID_output

```

```{r Chunk 4 Visualization: Fig 3a. Pear psylla adults summary}

#Adults graph
Fig3a <- ggplot(CID_output_adults, aes(x = Treatment, y = seasonalmean)) +
  geom_point(aes(group = Block), color = "grey") +
  geom_line(aes(group = Block), color = "grey") +
  stat_summary(fun = mean, colour = "black", geom = "point") +
  geom_text(data = data.frame(x = c("Ctrl", "R. fab.","M. film"), y = 10, Season = "Pre-installation", label = "a"), 
                   aes(x = x, y = y, label = label), size = 3) +
  geom_text(data = data.frame(x = c("Ctrl", "R. fab.","M. film"), y = c(2.9,2.8,1.9), Season = "Spring", label = c("a","a","b")), 
                   aes(x = x, y = y, label = label), size = 3) +
  scale_y_continuous(limits = c(0,NA)) +
  facet_wrap(~Season, scales = "free_y") +
  ylab('Avg. adults') +
  theme(strip.text = element_text(size = 10, hjust = 0),
          axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

#summary with SEM for table
Data_taps_analyzed <- do.call(data.frame, 
                          aggregate(seasonalmean~Season+Treatment, 
                                    data = CID_output_adults, 
                                    function(x) 
                                    c(mean = mean(x), 
                                      count = length(x),
                                      SEM = sd(x)/sqrt(length(x)))))

```

```{r Chunk 5 Statistics:    Pear psylla adult pre-installation}

#METHOD FOR PUBLICATION: 3b. summary data, log transform, fixed block. P = 0.16

#method 1a. summary data, no transform, random block. does not fit well
Data_test <- subset(CID_output_adults, Season == "Pre-installation")
model_glmm <- glmmTMB(formula = (seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmm))
print(Anova(model_glmm))
print(emmeans(model_glmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 1b. summary data, log transform, random block. P = 0.056
Data_test <- subset(CID_output_adults, Season == "Pre-installation")
model_glmm <- glmmTMB(formula = log(seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmm))
print(Anova(model_glmm))
print(emmeans(model_glmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 2. summary data, log transform, random block. P = 0.16
Data_test <- subset(CID_output_adults, Season == "Pre-installation")
model_nlme <- lme(log(seasonalmean) ~ Treatment, random = ~1 | Block, data = Data_test, method = "REML")
anova(model_nlme)
plot(model_nlme)
qqnorm(resid(model_nlme))
print(emmeans(model_nlme, list(pairwise ~ Treatment), adjust = "tukey"))

#method 3a. summary data, no transform, fixed block. P = 0.99
Data_test <- subset(CID_output_adults, Season == "Pre-installation")
model_conventional <- aov((seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 3b. summary data, log transform, fixed block. P = 0.16
Data_test <- subset(CID_output_adults, Season == "Pre-installation")
model_conventional <- aov(log(seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 4. no summary, nbinom 2, random plot nested in block. P = 0.22
Data_test <- subset(Data, Phase == "Pretreatment")
model_glmm_randomeffects <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), family = 'nbinom2', data = Data_test)
plot(simulateResiduals(model_glmm_randomeffects))
print(Anova(model_glmm_randomeffects))
print(emmeans(model_glmm_randomeffects, list(pairwise ~ Trt), adjust = "tukey"))

  
```

```{r Chunk 6 Statistics:    Pear psylla adult spring}

#METHOD FOR PUBLICATION: 3b. summary data, log transform, fixed block. P = 0.048

#method 1a. summary data, no transform, random block. P = 0.063
Data_test <- subset(CID_output_adults, Season == "Spring")
model_glmm <- glmmTMB(formula = seasonalmean ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmm))
print(Anova(model_glmm))
print(emmeans(model_glmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 1b. summary data, log transform, random block. P = 0.0033; Control greater than ref fab and metal film
Data_test <- subset(CID_output_adults, Season == "Spring")
model_glmm <- glmmTMB(formula = log(seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmm))
print(Anova(model_glmm))
print(emmeans(model_glmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 2. summary data, log transform, random block. P = 0.048
Data_test <- subset(CID_output_adults, Season == "Spring")
model_nlme <- lme(log(seasonalmean) ~ Treatment, random = ~1 | Block, data = Data_test, method = "REML")
anova(model_nlme)
plot(model_nlme)
qqnorm(resid(model_nlme))
print(emmeans(model_nlme, list(pairwise ~ Treatment), adjust = "tukey"))

#method 3a. summary data, no transform, fixed block. P = 0.17
Data_test <- subset(CID_output_adults, Season == "Spring")
model_conventional <- aov((seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 3b. summary data, log transform, fixed block. P = 0.048, control marginally greater than m. film
Data_test <- subset(CID_output_adults, Season == "Spring")
model_conventional <- aov(log(seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 4. no summary, nbinom 1, random plot nested in block. P = 0.028
Data_test <- Data_spring_all
Data_test <- subset(Data_test, Phase != 'Pretreatment')
model_glmm_randomeffects <- glmmTMB(formula = Adults ~ Trt + (1|Site/Trt), family = 'nbinom1', data = Data_test)
plot(simulateResiduals(model_glmm_randomeffects))
  print(Anova(model_glmm_randomeffects))
  print(emmeans(model_glmm_randomeffects, list(pairwise ~ Trt), adjust = "tukey"))

```

```{r Chunk 7 Summarization: Calculate pear psylla egg insect-days}
#First, summarize the multiple counts per plot into one number per plot per site per visit
Data_eggs <- do.call(data.frame, 
             aggregate(Eggs~Phase+Date+day+Site+Trt, 
                       data = Data, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
#Calculate cumulative insect days - eggs
#Build blank data frame
CID_output <- data.frame(Season=character(), 
                         Block=character(), 
                         Treatment=character(), 
                         CID=numeric(), 
                         Days_sum=numeric())
Data_eggs$ID <- paste(Data_eggs$Phase,Data_eggs$Site,Data_eggs$Trt,sep = "_")  #Make a unique identifer for each experimental unit by Period of study
IDs <- unique(Data_eggs$ID)                                   #Get a list of the IDs
for (value in IDs) {                                          #loop for each ID
  Data_subset <- subset(Data_eggs, ID == value)
  row.names(Data_subset) <- NULL
  for (row in 1:nrow(Data_subset)) {                          #loop to calculate within each ID
    if(row == 1){
      Insect_avg <- NA
      Days_difference <- NA
      Insectdays <- 0
      Days <- 0
      output <- data.frame(Insectdays,Days)
    }
    #pre-sample was only one date, so use the point data instead of trying to calculate CIDs
    if(unique(Data_subset$Phase == "Pretreatment")){
      Days <- 1
      Insectdays <- Data_subset[row,"Eggs.mean"]
      output <- data.frame(Insectdays,Days)
    }
    if(row > 1){
    Insect_current <- Data_subset[row,"Eggs.mean"]
    Insect_previous <- Data_subset[(row-1),"Eggs.mean"]
    Insect_avg <- (Insect_current + Insect_previous) / 2
    Day_current <- as.numeric(Data_subset[row,"day"])
    Day_previous <- as.numeric(Data_subset[row-1,"day"])
    if(Day_current < Day_previous){                             #check for out-of-order dates
      print("Error: Days not sorted chronologically")
      print(value)
    }
    Days <- Day_current - Day_previous
    Insectdays <- Insect_avg*Days
    newrow <- data.frame(Insectdays,Days)
    output <- rbind(output, newrow)
    }
  }
  Season <- unique(Data_subset$Phase)
  Block <- unique(Data_subset$Site)
  Treatment <- unique(Data_subset$Trt)
  CID <- sum(output$Insectdays)
  Days_sum <- sum(output$Days)
  CID_newrow <- data.frame(Season,Block,Treatment, CID, Days_sum)
  CID_output <- rbind(CID_output, CID_newrow)
}

#Don't want want to analyze summer data due to reflective materials being removed at different times between sites
CID_output <- subset(CID_output, Season != 'Leaves')
CID_output$Season <- factor(CID_output$Season,
                    levels=c("Pretreatment","Buds"),
                    labels=c("Pre-installation", "Spring"))
CID_output$Treatment <- factor(CID_output$Treatment,
                    levels=c("Control","Extenday","Mylar"),
                    labels=c("Ctrl", "R. fab.","M. film"))
CID_output_eggs <- CID_output
CID_output_eggs$seasonalmean <- (CID_output$CID / CID_output$Days_sum)

```

```{r Chunk 8 Visualization: Fig. 3b. Pear psylla eggs summary}

Fig3b <- ggplot(CID_output_eggs, aes(x = Treatment, y = seasonalmean)) +
  geom_point(aes(group = Block), color = "grey") +
  geom_line(aes(group = Block), color = "grey") +
  stat_summary(fun = mean, colour = "black", geom = "point") +
  geom_text(data = data.frame(x = c("Ctrl", "R. fab.","M. film"), y = c(9.5,7,10), Season = "Pre-installation", label = c("a","b","a")), 
                   aes(x = x, y = y, label = label), size = 3) +
  geom_text(data = data.frame(x = c("Ctrl", "R. fab.","M. film"), y = c(45,45,40), Season = "Spring", label = "a"), 
                   aes(x = x, y = y, label = label), size = 3) +
  scale_y_continuous(limits = c(0,NA)) +
  ylab('Avg. eggs') +
  facet_wrap(~Season, scales = "free_y") +
  theme(strip.text = element_text(size = 10, hjust = 0),
          axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

#summary with SEM for table
Data_eggs_analyzed <- do.call(data.frame, 
                          aggregate(CID/Days_sum~Season+Treatment, 
                                    data = CID_output_eggs, 
                                    function(x) 
                                    c(mean = mean(x), 
                                      count = length(x),
                                      SEM = sd(x)/sqrt(length(x)))))

#Figure 3 output
  Fig3 <- (ggarrange(
    Fig3a,
    Fig3b,
    ncol = 2,
    labels = c("A", "B"),
    align="hv",common.legend = TRUE
    ))
  
ggsave(plot = Fig3, file = "Fig3.png", width = 17, height = 5, unit = "cm")

```

```{r Chunk 9 Statistics:    Pear psylla egg pre-installation}

#METHOD FOR PUBLICATION: 3b. summary data, log transform, fixed block.

#method 1a. summary data, no transform, random block. P = 0.00008. r. fab lowest
Data_test <- subset(CID_output_eggs, Season == "Pre-installation")
model_glmmm <- glmmTMB(formula = seasonalmean ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmmm))
print(Anova(model_glmmm))
print(emmeans(model_glmmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 1b. summary data, log transform, random block. P = 0.006. r. fab lowest
Data_test <- subset(CID_output_eggs, Season == "Pre-installation")
model_glmmm <- glmmTMB(formula = log(seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmmm))
print(Anova(model_glmmm))
print(emmeans(model_glmmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 2. summary data, log transform, random block. P = 0.026
Data_test <- subset(CID_output_eggs, Season == "Pre-installation")
model_nlme <- lme(log(seasonalmean) ~ Treatment, random = ~1 | Block, data = Data_test, method = "REML")
anova(model_nlme)
plot(model_nlme)
qqnorm(resid(model_nlme))
print(emmeans(model_nlme, list(pairwise ~ Treatment), adjust = "tukey"))

#method 3a. summary data, no transform, fixed block. P = 0.029. r. fab less that m. film; r. fab marginally less than ctrl
Data_test <- subset(CID_output_eggs, Season == "Pre-installation")
model_conventional <- aov(seasonalmean ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 3b. summary data, log transform, fixed block. P = 0.026. r. fab less that m. film & ctrl
Data_test <- subset(CID_output_eggs, Season == "Pre-installation")
model_conventional <- aov(log(seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 4. no summary, nbinom 2, random plot nested in block. P = 0.022. r. fab less that m. film; r. fab marginally less than ctrl
Data_test <- Data_spring_all
Data_test <- subset(Data_test, Phase == 'Pretreatment')
model_glmm_randomeffects <- glmmTMB(formula = Eggs ~ Trt + (1|Site/Trt), family = 'nbinom1', data = Data_test)
plot(simulateResiduals(model_glmm_randomeffects))
print(Anova(model_glmm_randomeffects))
print(emmeans(model_glmm_randomeffects, list(pairwise ~ Trt), adjust = "tukey"))

```

```{r Chunk 10 Statistics:   Pear psylla egg spring}  

#METHOD FOR PUBLICATION: 3b. summary data, log transform, fixed block. P = 0.23

#method 1a. summary data, no transform, random block. P = 0.27
Data_test <- subset(CID_output_eggs, Season == "Spring")
model_glmmm <- glmmTMB(formula = (seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmmm))
print("Pear psylla eggs")
print(Anova(model_glmmm))
print(emmeans(model_glmmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 1b. summary data, log transform, random block. P = 0.11
Data_test <- subset(CID_output_eggs, Season == "Spring")
model_glmmm <- glmmTMB(formula = log(seasonalmean) ~ Treatment + (1|Block), data = Data_test)
plot(simulateResiduals(model_glmmm))
print("Pear psylla eggs")
print(Anova(model_glmmm))
print(emmeans(model_glmmm, list(pairwise ~ Treatment), adjust = "tukey"))

#method 2. summary data, log transform, random block. P = 0.23
Data_test <- subset(CID_output_eggs, Season == "Spring")
model_nlme <- lme(log(seasonalmean) ~ Treatment, random = ~1 | Block, data = Data_test, method = "REML")
anova(model_nlme)
plot(model_nlme)
qqnorm(resid(model_nlme))
print(emmeans(model_nlme, list(pairwise ~ Treatment), adjust = "tukey"))

#method 3a. summary data, no transform, fixed block. P = 0.40
Data_test <- subset(CID_output_eggs, Season == "Spring")
model_conventional <- aov((seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 3b. summary data, log transform, fixed block. P = 0.23
Data_test <- subset(CID_output_eggs, Season == "Spring")
model_conventional <- aov(log(seasonalmean) ~ Treatment + Block, data = Data_test)
summary(model_conventional)
print(emmeans(model_conventional, list(pairwise ~ Treatment), adjust = "tukey"))
plot(model_conventional, 1)
plot(model_conventional, 2)

#method 4. no summary, nbinom 2, random plot nested in block. P = 0.48
Data_test <- Data_spring_all
Data_test <- subset(Data_test, Phase != 'Pretreatment')
model_glmm_randomeffects <- glmmTMB(formula = Eggs ~ Trt + (1|Site/Trt), family = 'nbinom1', data = Data_test)
plot(simulateResiduals(model_glmm_randomeffects))
print(Anova(model_glmm_randomeffects))
print(emmeans(model_glmm_randomeffects, list(pairwise ~ Trt), adjust = "tukey"))
  


```

```{r Chunk 11 Supplementary analysis and diagnostics}

#summarize nymph counts in spring to check how abundant they were
#reload data
Data <- read_excel("2003 - data - beats and buds.xlsx",
                   skip = 9,
                   guess_max = 10000) #guessmax prevents columns with many initial banks being imported as a logical FALSE for not NAs
Data <- subset(Data, day < 200) #Exclude the late sample in August since it was over a month divorced from previous samples
Data <- subset(Data, !(Site == "E" & day == 77)) #exclude the early pre-treatment sample at site E so we only analyze the one sample closer to groundcover installation
Data_spring_post <- subset(Data, Phase == 'Buds')

#Summary small nymphs; they were rare in spring (e.g., not found at site E, max of 0.03 per leaf at site D):
Data_smalls <- do.call(data.frame, 
             aggregate(Small~day+Site+Trt, 
                       data = Data_spring_post, 
                       function(x) 
                       c(mean = mean(x), 
                       count = length(x))))
Data_smalls <- do.call(data.frame, 
             aggregate(Small.mean~Site, 
                       data = Data_smalls, 
                       function(x) 
                       c(max = max(x), 
                       count = length(x))))



```


